\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-09-21 notes/2loopNLSM.tex
fixed all n vs N vs mathcalN and YZ + J

The final constraints comes from ensuring that the ansatz correctly
reproduces the generalized unitarity cuts of the desired theory, thus
guaranteeing that the numerators encode the correct physical,
theory-dependent information.  Generalized unitarity methods has been
reviewed in many places so we will be light on details here
\cite{BCFUnitarity, Bern:2015ooa, Bourjaily:2017wjl,
  JJHenrikReview, FivePointN4BCJ, Bern:2012uf, BCJreview,
  Edison:2022jln, Edison:2022smn}, which was recently optimized for effective field theories like NLSM \cite{Carrasco:2023qgz}. Generalized unitarity equates products of on-shell tree amplitudes,
encoded via a graph $\gamma$, to sums over compatible\footnote{A
  diagram $g$ is compatible with the cut diagram $\gamma$ if $g$ is
  isomorphic to $\gamma$ or to any of the additional factorization
  channels of $\gamma$. } diagram numerators evaluated on the support
of the on-shell conditions and weighted by their uncut propagators
\begin{equation}
\label{eq:SchematicCutEquation}
\text{Cut}[\gamma]= \sum_{\substack{\text{states}\\\text{crossing } E(\gamma)}} \prod_{v \in V(\gamma)} \atree(v)
= \sum_{g \text{ compat } \gamma}
\frac{n[g] \vert_\text{cut}}{\prod \limits_{\text{uncut}} \ell^2}
 \, ,
\end{equation}
where $V(\gamma)$ are the vertices of $\gamma$ and $E(\gamma)$ its
edges.  The sum over states hides much of the complexity of cut
construction: for scalars the process is trivial while spinning states
greatly complicate matters (see
Refs.~\cite{Kosmopoulos:2020pcd,Edison:2022jln} for recent discussion
and detailed examples of evaluating state sums).  As an example of the
two types of expressions appearing in \cref{eq:SchematicCutEquation},
consider the so-called iterated two-particle cut which can be
represented as a product of trees via
\begin{equation}
  \label{eq:example-cut-prod}
  \ExCutLabeled{1}{2}{3}{4} = \!\!\!\!\! \sum_{\substack{\text{states of }\\ \{\ell_1, \ell_2, \ell_3, \ell_4\}}} \!\!\!\!\!\!\!
  \atree(1,2,\ell_1,\ell_2) \atree(-\ell_1,-\ell_2,\ell_3,\ell_4) \atree(-\ell_3,-\ell_4,3,4)
\end{equation}
or as a collection of diagram numerators via
\begin{align}
  \label{eq:example-cut-sum}
  \ExCutLabeled{1}{2}{3}{4} &=
  \frac{\delta(\textcolor{cut-color}{\ell_1^2})
    \delta(\textcolor{cut-color}{\ell_2^2})
    \delta(\textcolor{cut-color}{\ell_3^2})
    \delta(\textcolor{cut-color}{\ell_4^2})}{(\textcolor{ucp-color}{\ell_5})^2
    (\textcolor{ucp-color}{\ell_6})^2
    (\textcolor{ucp-color}{\ell_7})^2}
  \ n\left[
    \IUCContrib{1}{4}{3}{2}
                              \right] \notag \\
  &\quad + 
  \frac{\delta(\textcolor{cut-color}{\ell_1^2})
    \delta(\textcolor{cut-color}{\ell_2^2})
    \delta(\textcolor{cut-color}{\ell_3^2})
    \delta(\textcolor{cut-color}{\ell_4^2})}
    {(\textcolor{ucp-color}{\ell_5})^2
    (\textcolor{ucp-color}{\ell_6})^2
    (\textcolor{ucp-color}{\ell_7'})^2}
  \ n\left[
    \IUCContrib{1}{3}{4}{2}
    \right] + 25\text{ others.}
\end{align}
When drawing cut diagrams, we will use gray blobs as in the left-hand
sides of \cref{eq:example-cut-prod,eq:example-cut-sum} and all drawn
legs are assumed to be on-shell.  When drawing numerator contributions
to cuts, we will not use blobs on vertices, and will use
\textcolor{cut-color}{dashed lines} bisecting edges to denote cut
lines and use colored edges to highlight the
\textcolor{ucp-color}{uncut propagators}.  From now on we will leave
the cut $\delta(\ell^2)$s implicit.  \jm{Again, I love the colors!
  Would it be possible to add $\ell$ labels to the RHS of
  \cref{eq:example-cut-sum} to show what $\ell_5$, $\ell_6$... are?}
\ace{It is doable, will work on it once our text is a bit more happy}

In color-charged theories, both sides of the cut can be further
decomposed according to the color algebra.  For theories charged under
the adjoint of $SU(N)$ via $f^{abc}$-dressed amplitudes, it is
convenient to project onto a specific element of the Del
Duca-Dixon-Maltoni (DDM) color basis \cite{DixonMaltoni} by only
inserting color-ordered amplitudes, a la the right hand side of
\cref{eq:ddm-basis} in the product of trees, and restricting the set
of compatible diagrams to those whose color factors reduce to the
appropriate DDM element when summing over \emph{only the uncut color
  contractions}.  In this situation, the color becomes an overall
factor on the entire cut equality and thus can be ignored.  Cuts that
are organized in this manner are known as ``color-ordered'' cuts, in
the same sense as color-ordered tree amplitudes.  Since the DDM basis
(and dual Kleiss-Kuijf amplitude basis \cite{Kleiss:1988ne}) have
$(n-2)!$ elements, one generally needs to evaluate
$(n_1-2)! (n_2-2)!...$ separate color-orderings of the same cut, where
$n_1$, $n_2$... are the multiplicities of the amplitudes making up the
cut. As an example, consider the ``bowtie'' cut topology -- one of the
factorization channels of the iterated unitarity cut -- which has two
topologically distinct color-ordered expansions
\begin{equation}
  \label{eq:kt-color-ord}
   \NMCutC =
  \begin{cases}
     s\text{-}t \text{ channel}
    &
      \frac{1}{\textcolor{ucp-color}{s_{12}}}
      n \left[\scalebox{0.65}{\KTBTContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      n \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]\\
    t\text{-}u \text{ channel}
    & 
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      n \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_2^2}} \
      n\left[
      \scalebox{0.65}{\KTDBContrib{1}{3}{4}{2}}
      \right]
  \end{cases}
\end{equation}
which serve as a complete basis for the full cut.

@@

\subsection{One-loop color-dual integrands}\label{oneLoopCK}
Our first application of this theory for color-dual construction at
loop level is for self-dual Yang-Mills (SDYM). At tree-level, the
off-shell cubic vertices can be written in terms of light cone
coordinates \cite{Monteiro2011pc}:
\begin{equation}
X(p,k) = p_u k_w-p_w k_u
\end{equation}
At loop-level, this construction needs to be analytically continued to
general dimension in order to apply dimensional regularization at
one-loop. This can be done using the cubic semi-YM vertices of the
previous section:
\begin{equation} \label{DdimSDYM}
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= \mathcal{X}(k_1,k_2) = (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3) -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3)
\end{equation}
Plugging in on-shell all-plus helicity states in light-cone gauge will
yield precisely the 4D SDYM vertex, up to an unphysical phase \ace{Can
  we provide an example?} \nhp{yes, provided}:
  \be
  \mathcal{X}(k^+_1,k^+_2) = \frac{\langle 12\rangle^3}{ \langle23\rangle \langle 31\rangle} \sim \langle 12\rangle = k_{1,u} k_{2,w}-k_{1,w} k_{2,u}
 \ee
where $\mathcal{X}(k^+_1,k^+_2) = A^{\text{YM}}(1^+,2^+,3^-)$ in light cone gauge, and in the second step applied momentum conservation to redefinition of the spinor bracket, $\langle 12\rangle \rightarrow X(p,k)$ defined in \cref{Conventions}. Of course, the form of \cref{DdimSDYM} has the advantage of
permitting a $D$-dimensional construction of the one-loop
integrand. Below we provide a four-point example: \nhp{change ordering on legs}
\begin{equation}
\simplebox = g_{\text{YM}}^4 \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\mathcal{X}(k_3,l_3)\mathcal{X}(k_4,l_4)\rangle
\end{equation}
where $l_i = l+k_1+k_2+\cdots+ k_i$ and the bracket $\langle \,\cdots\rangle$ indicates that we have applied the gauge fixed state projector, $\sum {\varepsilon^\mu_{(+l_i)}\varepsilon^\nu_{(-l_i)}}=\eta^{\mu\nu}$ on all internal polarizations. In general, the $n$-gon diagram would be,
\begin{equation}
\mathcal{N}^{\text{SDYM}}_{n\text{-gon}}= \nGonVector = g_{\text{YM}}^n \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\cdots \mathcal{X}(k_n,l_n) \rangle
\end{equation}
Due to the state-sum of internal loop factors,
$\sum {\varepsilon_{(+l)}\varepsilon_{(-l)}}\sim D$, the integrand above
depends explicitly on the spacetime dimension, $D$. By taking a
derivative\footnote{Another way to understand this derivative is that
  it selects the large $D$ behavior of the integrand.}, we can recover
the integrand needed for all-plus one-loop amplitudes with an internal
scalar:
\begin{equation}
\partial_D \mathcal{N}^{\text{SDYM}}_{n\text{-gon}}=\nGonScalar  = (2 g_{\text{YM}})^n (l_1 \varepsilon_1)(l_2 \varepsilon_2)\cdots (l_n \varepsilon_n)
\end{equation}
This integrand numerator with internal scalar loop is precisely what
one would obtain from the Feynman rules of $YZ$-theory, absent the $\bar{Z}Z$
internal vector loop. For more background, we refer the reader to
\cref{sec:CKLagrangians}. When plugging on the on-shell states of $YZ$-theory, we thus obtain the following expression for the NLSM one-loop $n$-gon numerator:
\begin{equation}
 \mathcal{N}^{\text{NLSM}}_{n\text{-gon}}=\left[\partial_D  \mathcal{N}^{\text{SDYM}}_{n\text{-gon}}\right]^{\epsilon \rightarrow k}=\nGonScalar =  f_\pi^{-n} [12][23]\cdots [n1]
\end{equation}
Where we have define the antisymmetric kinematic variable, $[ij] = l_i^2 - l_j^2$, since when plugging on longitudinal modes, the $n$-gon can be re-expressed completely in terms of inverse propagators, $2(k_i \cdot l_i) = l_{i}^2-l^2_{i-1}$. We have verified through 10-point one-loop that this $n$-gon expression is a valid color-dual representation for NLSM.

@@

Ignoring tadpoles and BEL graphs, there are 14 cubic four-point two-loop
graphs (see \cref{fig:MaxCuts}) which are related to each other via 21
Jacobi relations.  As mentioned in \cref{sec:bootstrap},
color-kinematics duality ensures that the numerator of every graph can
be expressed in terms of a basis of the double box and penta-triangle,
\begin{equation}
\label{eq:jacobi-basis}
\NLSMBasisDoubleBox \text{ and } \NLSMBasisPentaTriangle .
\end{equation}
For a scalar theory like NLSM the numerator only depends on dot
products of momenta.  The basis of momentum invariants for both of
these graphs can be taken to be
\begin{equation}
\label{eq:scalar-basis}
\{ k_{13}, k_{23}, k_{15}, k_{16}, k_{25}, k_{26}, k_{35}, k_{36}, k_{55}, k_{56}, k_{66} \},
\end{equation}
where $k_{ij}$ means $k_i \cdot k_j$, $k_5=\ell_1$, and $k_6 = \ell_2$.
%%%%
% Basis for both diagrams
% {d[p[1], p[3]], d[p[1], p[6]], d[p[1], p[7]], d[p[2], p[3]],  d[p[2], p[6]], d[p[2], p[7]], d[p[3], p[6]], d[p[3], p[7]],  d[p[6], p[6]], d[p[6], p[7]], d[p[7], p[7]]}
%%%%
Every vertex in NLSM scales as $k^2$ so the numerator of each diagram
scales as $k^{12}$.  At this mass dimension there are 8,008
independent dot products of momenta.  With two basis graphs there is a
total of 16,016 ansatz parameters.\footnote{\texttt{Mathematica} is
  not suited to equations with sixteen thousand parameters so a
  different method is desirable.  One possibility is to use a solver
  designed for integral reduction.  However, these solvers typically
  assume that the matrix has a very low density which is not the case
  for equations coming from NLSM which typically have a density of a
  few percent.  While the equations in YM naturally separate into
  different ``helicity'' sectors (or $e_i \cdot e_j$ and
  $e_i \cdot k_j$ sectors in $D$ dimensions), there is no such
  separation of sectors in a scalar theory where any product of
  momenta can interfere with any other.  The NLSM constraint equations
  were solved using a custom solver, \texttt{FiniteFieldSolve},
  designed to solve large linear systems of arbitrary density.
  \texttt{FiniteFieldSolve} will be released publicly shortly.\ace{I
    want to rephrase this slightly, but need to think about it a bit
    more}} Imposing the boundary Jacobi relations eliminates all but
4,473 of the original free parameters.  Further imposing graph
symmetries reduces the number of free parameters to 1,243.

@@

The final essential constraint is unitarity.  The two physical cuts
that need to be enforced, shown in \cref{fig:emu}, only involve
four-point vertices.  Notably this means that even at two loops,
unitarity is still only probing the on-shell four-point NLSM amplitude.
Unitarity cuts besides those shown in \cref{fig:emu} are either
pathological or vanish because they involve a 3pt point on-shell
amplitude (which vanishes on general grounds for scalar theories
\ace{This feels like overstating things.  For instance, what about
  $\phi^3$ or BAS?}).  In order to manifest the $\mathbb{Z}_2$
symmetry of NLSM at the integrand level, we require that all
non-pathological cuts containing a 3pt vertex (such as the maximal
cut) vanishs.  Imposing all relevant cuts leaves 765 free parameters
representing the generalized gauge freedom in the answer.

% It is possible to automatically enforce (or ``bake in'') the double
% box and penta-triangle max cuts on the ansatz by writing the ansatz
% in terms of inverse propagators (and remaining irreducible scalar
% products) and then only keeping terms that involve at least one
% inverse propagator since those terms will vanish on those max cuts,
% just as the product of 3pt scalar amplitudes will.  However, for the
% sake of simplicity we do not take this route.

@@

\ace{Lets tighten up this paragraph a bit, especially the second half}
It is tempting to try to match the pion four-point two-loop integrand to ZM
on the cuts, but given that the cuts only probe the on-shell four-point
amplitude, which is either subtle or vanishes for ZM, it is better to
compare the off-shell numerators directly.  As local functions, the
numerators never suffer from any of the subtleties of on-shell 2D
kinematics.  Since the physical cuts are difficult to compare, ZM
theory is being used here simply as a tool to eliminate generalized
gauge freedom; ZM theory has the right power counting and kinematic
structure in 2D to compare against.  Since ZM theory is only being
used for the kinematical structure of its numerators, it is natural to
introduce a proportionality constant $z$ between the NLSM and ZM
numerators $n_\text{NLSM} \vert_{2D} = z ~ n_\text{ZM}$.  The
parameter $z$, which would normally be fixed on cuts, effectively
encodes the discrepancy between the coupling constants of the two
theories.\footnote{Floating the proportionality parameter $z$ is also
  motivated because ZM is believed to be quantum mechanically
  different from NLSM even though they are dual classically in two
  dimensions \cite{Nappi:1979ig}.}  This introduces an extra degree of
freedom into the problem.  Mechanically, the pion numerator is matched
to ZM by first taking every possible assignment of right and left
movers for the external particles and then restricting the pion
numerator to 2D.  The loop momenta are restricted to 2D but left
off-shell.  After performing the 2D matching, there are only 365
parameters of generalized gauge freedom.

@@

Much of the complexity for $D$-dimensional vector theories at
loop-level arises due to the mixing between external polarization and
internal loop momenta. As noted, one advantage of expanding ansatz
construction towards functions of rational kinematics is the prospect
of stripping out overall factors of gauge invariant tensors. Rather
than building an ansatz from the irreducible scalar products of
\cref{eq:ym-basis},
\begin{equation}
\mathcal{N}^{\text{YM}}_{m,L} = \sum_i \mathcal{O}_i \mathcal{R}^{(i)}_{m,L}
\end{equation}
where $\mathcal{O}_i$ are the basis tensors of
\cite{Bern:2017tuc,Carrasco:2019yyn}, and $ \mathcal{R}_i$ is a
rational function of scalar products.

>>>>

The final constraints comes from ensuring that the ansatz correctly
reproduces the generalized unitarity cuts of the desired theory, thus
guaranteeing that the numerators encode the correct physical,
theory-dependent information.  Generalized unitarity methods has been
reviewed in many places so we will be light on details here
\cite{BCFUnitarity, Bern:2015ooa, Bourjaily:2017wjl,
  JJHenrikReview, FivePointN4BCJ, Bern:2012uf, BCJreview,
  Edison:2022jln, Edison:2022smn}, which was recently optimized for effective field theories like NLSM \cite{Carrasco:2023qgz}. Generalized unitarity equates products of on-shell tree amplitudes,
encoded via a graph $\gamma$, to sums over compatible\footnote{A
  diagram $g$ is compatible with the cut diagram $\gamma$ if $g$ is
  isomorphic to $\gamma$ or to any of the additional factorization
  channels of $\gamma$. } diagram numerators evaluated on the support
of the on-shell conditions and weighted by their uncut propagators
\begin{equation}
\label{eq:SchematicCutEquation}
\text{Cut}[\gamma]= \sum_{\substack{\text{states}\\\text{crossing } E(\gamma)}} \prod_{v \in V(\gamma)} \atree(v)
= \sum_{g \text{ compat } \gamma}
\frac{N[g] \vert_\text{cut}}{\prod \limits_{\text{uncut}} \ell^2}
 \, ,
\end{equation}
where $V(\gamma)$ are the vertices of $\gamma$ and $E(\gamma)$ its
edges.  The sum over states hides much of the complexity of cut
construction: for scalars the process is trivial while spinning states
greatly complicate matters (see
Refs.~\cite{Kosmopoulos:2020pcd,Edison:2022jln} for recent discussion
and detailed examples of evaluating state sums).  As an example of the
two types of expressions appearing in \cref{eq:SchematicCutEquation},
consider the so-called iterated two-particle cut which can be
represented as a product of trees via
\begin{equation}
  \label{eq:example-cut-prod}
  \ExCutLabeled{1}{2}{3}{4} = \!\!\!\!\! \sum_{\substack{\text{states of }\\ \{\ell_1, \ell_2, \ell_3, \ell_4\}}} \!\!\!\!\!\!\!
  \atree(1,2,\ell_1,\ell_2) \atree(-\ell_1,-\ell_2,\ell_3,\ell_4) \atree(-\ell_3,-\ell_4,3,4)
\end{equation}
or as a collection of diagram numerators via
\begin{align}
  \label{eq:example-cut-sum}
  \ExCutLabeled{1}{2}{3}{4} &=
  \frac{\delta(\textcolor{cut-color}{\ell_1^2})
    \delta(\textcolor{cut-color}{\ell_2^2})
    \delta(\textcolor{cut-color}{\ell_3^2})
    \delta(\textcolor{cut-color}{\ell_4^2})}{(\textcolor{ucp-color}{\ell_5})^2
    (\textcolor{ucp-color}{\ell_6})^2
    (\textcolor{ucp-color}{\ell_7})^2}
  \ N\left[
    \IUCContrib{1}{4}{3}{2}
                              \right] \notag \\
  &\quad + 
  \frac{\delta(\textcolor{cut-color}{\ell_1^2})
    \delta(\textcolor{cut-color}{\ell_2^2})
    \delta(\textcolor{cut-color}{\ell_3^2})
    \delta(\textcolor{cut-color}{\ell_4^2})}
    {(\textcolor{ucp-color}{\ell_5})^2
    (\textcolor{ucp-color}{\ell_6})^2
    (\textcolor{ucp-color}{\ell_7'})^2}
  \ N\left[
    \IUCContrib{1}{3}{4}{2}
    \right] + 25\text{ others.}
\end{align}
When drawing cut diagrams, we will use gray blobs as in the left-hand
sides of \cref{eq:example-cut-prod,eq:example-cut-sum} and all drawn
legs are assumed to be on-shell.  When drawing numerator contributions
to cuts, we will not use blobs on vertices, and will use
\textcolor{cut-color}{dashed lines} bisecting edges to denote cut
lines and use colored edges to highlight the
\textcolor{ucp-color}{uncut propagators}.  From now on we will leave
the cut $\delta(\ell^2)$s implicit.

%% \jm{Again, I love the colors!
%%  Would it be possible to add $\ell$ labels to the RHS of
%5  \cref{eq:example-cut-sum} to show what $\ell_5$, $\ell_6$... are?}
%%\ace{It is doable, will work on it once our text is a bit more happy}

In color-charged theories, both sides of the cut can be further
decomposed according to the color algebra.  For theories charged under
the adjoint of $SU(N)$ via $f^{abc}$-dressed amplitudes, it is
convenient to project onto a specific element of the Del
Duca-Dixon-Maltoni (DDM) color basis \cite{DixonMaltoni} by only
inserting color-ordered amplitudes, a la the right hand side of
\cref{eq:ddm-basis} in the product of trees, and restricting the set
of compatible diagrams to those whose color factors reduce to the
appropriate DDM element when summing over \emph{only the uncut color
  contractions}.  In this situation, the color becomes an overall
factor on the entire cut equality and thus can be ignored.  Cuts that
are organized in this manner are known as ``color-ordered'' cuts, in
the same sense as color-ordered tree amplitudes.  Since the DDM basis
(and dual Kleiss-Kuijf amplitude basis \cite{Kleiss:1988ne}) have
$(n-2)!$ elements, one generally needs to evaluate
$(n_1-2)! (n_2-2)!...$ separate color-orderings of the same cut, where
$n_1$, $n_2$... are the multiplicities of the amplitudes making up the
cut. As an example, consider the ``bowtie'' cut topology -- one of the
factorization channels of the iterated unitarity cut -- which has two
topologically distinct color-ordered expansions
\begin{equation}
  \label{eq:kt-color-ord}
   \NMCutC =
  \begin{cases}
     s\text{-}t \text{ channel}
    &
      \frac{1}{\textcolor{ucp-color}{s_{12}}}
      N \left[\scalebox{0.65}{\KTBTContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      N \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]\\
    t\text{-}u \text{ channel}
    & 
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      N \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_2^2}} \
      N\left[
      \scalebox{0.65}{\KTDBContrib{1}{3}{4}{2}}
      \right]
  \end{cases}
\end{equation}
which serve as a complete basis for the full cut.

@@

\subsection{One-loop color-dual integrands}\label{oneLoopCK}
Our first application of this theory for color-dual construction at
loop level is for self-dual Yang-Mills (SDYM). At tree-level, the
off-shell cubic vertices can be written in terms of light cone
coordinates \cite{Monteiro2011pc}:
\begin{equation}
X(p,k) = p_u k_w-p_w k_u
\end{equation}
At loop-level, this construction needs to be analytically continued to
general dimension in order to apply dimensional regularization at
one-loop. This can be done using the cubic semi-YM vertices of the
previous section:
\begin{equation} \label{DdimSDYM}
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= \mathcal{X}(k_1,k_2) = (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3) -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3)
\end{equation}
Plugging in on-shell all-plus helicity states in light-cone gauge will
yield precisely the 4D SDYM vertex, up to an unphysical phase \ace{Can
  we provide an example?} \nhp{yes, provided}:
  \be
  \mathcal{X}(k^+_1,k^+_2) = \frac{\langle 12\rangle^3}{ \langle23\rangle \langle 31\rangle} \sim \langle 12\rangle = k_{1,u} k_{2,w}-k_{1,w} k_{2,u}
 \ee
where $\mathcal{X}(k^+_1,k^+_2) = A^{\text{YM}}(1^+,2^+,3^-)$ in light cone gauge, and in the second step applied momentum conservation to redefinition of the spinor bracket, $\langle 12\rangle \rightarrow X(p,k)$ defined in \cref{Conventions}. Of course, the form of \cref{DdimSDYM} has the advantage of
permitting a $D$-dimensional construction of the one-loop
integrand. Below we provide a four-point example: \nhp{change ordering on legs}
\begin{equation}
\simplebox = g_{\text{YM}}^4 \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\mathcal{X}(k_3,l_3)\mathcal{X}(k_4,l_4)\rangle
\end{equation}
where $l_i = l+k_1+k_2+\cdots+ k_i$ and the bracket $\langle \,\cdots\rangle$ indicates that we have applied the gauge fixed state projector, $\sum {\varepsilon^\mu_{(+l_i)}\varepsilon^\nu_{(-l_i)}}=\eta^{\mu\nu}$ on all internal polarizations. In general, the $n$-gon diagram would be,
\begin{equation}
N^{\text{SDYM}}_{n\text{-gon}}= \nGonVector = g_{\text{YM}}^n \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\cdots \mathcal{X}(k_n,l_n) \rangle
\end{equation}
Due to the state-sum of internal loop factors,
$\sum {\varepsilon_{(+l)}\varepsilon_{(-l)}}\sim D$, the integrand above
depends explicitly on the spacetime dimension, $D$. By taking a
derivative\footnote{Another way to understand this derivative is that
  it selects the large $D$ behavior of the integrand.}, we can recover
the integrand needed for all-plus one-loop amplitudes with an internal
scalar:
\begin{equation}
\partial_D N^{\text{SDYM}}_{n\text{-gon}}=\nGonScalar  = (2 g_{\text{YM}})^n (l_1 \varepsilon_1)(l_2 \varepsilon_2)\cdots (l_n \varepsilon_n)
\end{equation}
This integrand numerator with internal scalar loop is precisely what
one would obtain from the Feynman rules of $YZ$-theory, absent the $\bar{Z}Z$
internal vector loop. For more background, we refer the reader to
\cref{sec:CKLagrangians}. When plugging on the on-shell states of $YZ$-theory, we thus obtain the following expression for the NLSM one-loop $n$-gon numerator:
\begin{equation}
N^{\text{NLSM}}_{n\text{-gon}}=\left[\partial_D  N^{\text{SDYM}}_{n\text{-gon}}\right]^{\epsilon \rightarrow k}=\nGonScalar =  f_\pi^{-n} [12][23]\cdots [n1]
\end{equation}
Where we have define the antisymmetric kinematic variable, $[ij] = l_i^2 - l_j^2$, since when plugging on longitudinal modes, the $n$-gon can be re-expressed completely in terms of inverse propagators, $2(k_i \cdot l_i) = l_{i}^2-l^2_{i-1}$. We have verified through 10-point one-loop that this $n$-gon expression is a valid color-dual representation for NLSM.

@@

Ignoring tadpoles and BEL graphs, there are 14 cubic four-point two-loop
graphs (see \cref{fig:MaxCuts}) which are related to each other via 21
Jacobi relations.  As mentioned in \cref{sec:bootstrap},
color-kinematics duality ensures that the numerator of every graph can
be expressed in terms of a basis of the double box and penta-triangle,
\begin{equation}
\label{eq:jacobi-basis}
\NLSMBasisDoubleBox \text{ and } \NLSMBasisPentaTriangle .
\end{equation}
For a scalar theory like NLSM the numerator only depends on dot
products of momenta.  The basis of momentum invariants for the numerators assigned to both of
these graphs can be taken to be
\begin{equation}
\label{eq:scalar-basis}
\{ k_{13}, k_{23}, k_{15}, k_{16}, k_{25}, k_{26}, k_{35}, k_{36}, k_{55}, k_{56}, k_{66} \},
\end{equation}
where $k_{ij}$ means $k_i \cdot k_j$, $k_5=\ell_1$, and $k_6 = \ell_2$.
%%%%
% Basis for both diagrams
% {d[p[1], p[3]], d[p[1], p[6]], d[p[1], p[7]], d[p[2], p[3]],  d[p[2], p[6]], d[p[2], p[7]], d[p[3], p[6]], d[p[3], p[7]],  d[p[6], p[6]], d[p[6], p[7]], d[p[7], p[7]]}
%%%%
Every vertex in NLSM scales as $k^2$ so the numerator of each diagram
scales as $k^{12}$.  At this mass dimension there are 8,008
independent dot products of momenta.  With two basis graphs there is a
total of 16,016 ansatz parameters.\footnote{\texttt{Mathematica} is
  not suited to equations with sixteen thousand parameters so a
  different method is desirable.  One possibility is to use a solver
  designed for integral reduction.  However, these solvers typically
  assume that the matrix has a very low density which is not the case
  for equations coming from NLSM which typically have a density of a
  few percent.  While the equations in YM naturally separate into
  different ``helicity'' sectors (or $e_i \cdot e_j$ and
  $e_i \cdot k_j$ sectors in $D$ dimensions), there is no such
  separation of sectors in a scalar theory where any product of
  momenta can interfere with any other.  The NLSM constraint equations
  were solved using a custom solver, \texttt{FiniteFieldSolve},
  designed to solve large linear systems of arbitrary density.
  \texttt{FiniteFieldSolve} will be released publicly shortly.\ace{I
    want to rephrase this slightly, but need to think about it a bit
    more}} Imposing the boundary Jacobi relations eliminates all but
4,473 of the original free parameters.  Further imposing graph
symmetries reduces the number of free parameters to 1,243.

@@

The final essential constraint is unitarity.  The two physical cuts
that need to be enforced, shown in \cref{fig:emu}, only involve
four-point vertices.  Notably this means that even at two loops,
unitarity is still only probing the on-shell four-point NLSM amplitude.
Unitarity cuts besides those shown in \cref{fig:emu} are either
pathological or vanish because they involve a three-point point on-shell
amplitude. In order to manifest the $\mathbb{Z}_2$
symmetry of NLSM at the integrand level, we require that all
non-pathological cuts containing a three-point vertex (such as the maximal
cut) vanishes.  Imposing all relevant cuts leaves 765 free parameters
representing the generalized gauge freedom in the answer.

% It is possible to automatically enforce (or ``bake in'') the double
% box and penta-triangle max cuts on the ansatz by writing the ansatz
% in terms of inverse propagators (and remaining irreducible scalar
% products) and then only keeping terms that involve at least one
% inverse propagator since those terms will vanish on those max cuts,
% just as the product of three-point scalar amplitudes will.  However, for the
% sake of simplicity we do not take this route.

@@

\ace{Lets tighten up this paragraph a bit, especially the second half}
It is tempting to try to match the pion four-point two-loop integrand to ZM
on the cuts, but given that the cuts only probe the on-shell four-point
amplitude, which is either subtle or vanishes for ZM, it is better to
compare the off-shell numerators directly.  As local functions, the
numerators never suffer from any of the subtleties of on-shell 2D
kinematics.  Since the physical cuts are difficult to compare, ZM
theory is being used here simply as a tool to eliminate generalized
gauge freedom; ZM theory has the right power counting and kinematic
structure in 2D to compare against.  Since ZM theory is only being
used for the kinematical structure of its numerators, it is natural to
introduce a proportionality constant $z$ between the NLSM and ZM
numerators $n_\text{NLSM} \vert_{2D} = z ~ n_\text{ZM}$.  The
parameter $z$, which would normally be fixed on cuts, effectively
encodes the discrepancy between the coupling constants of the two
theories.\footnote{Floating the proportionality parameter $z$ is also
  motivated because ZM is believed to be quantum mechanically
  different from NLSM even though they are dual classically in two
  dimensions \cite{Nappi:1979ig}.}  This introduces an extra degree of
freedom into the problem.  Mechanically, the pion numerator is matched
to ZM by first taking every possible assignment of right and left
movers for the external particles and then restricting the pion
numerator to 2D.  The loop momenta are restricted to 2D but left
off-shell.  After performing the 2D matching to the following ZM basis numerators:
\be
\ZMBasisDoubleBox =\left[\begin{split}
  \langle (l_1+k_4) l_1 \rangle\langle l_1l_2 \rangle\langle l_2 (l_2+k_1) \rangle  \langle (l_2+k_1)(l_2+k_1+k_2) \rangle 
\\
\langle (l_1+k_3+k_4)(l_2+k_1+k_2) \rangle\langle (l_1+k_3+k_4)(l_1+k_4) \rangle
\end{split}\right]
\ee
\be
\ZMBasisPentaTriangle =\left[\begin{split}
  \langle (l_1+k_4) l_1 \rangle\langle l_1l_2 \rangle\langle l_2 (l_2+k_1) \rangle  \langle (l_2+k_1)(l_1-k_1) \rangle 
\\
\langle (l_1+k_3+k_4)(l_1-k_1) \rangle\langle (l_1+k_3+k_4)(l_1+k_4) \rangle
\end{split}\right]
\ee
there are only 365 parameters of generalized gauge freedom. An overview of ZM theory and its color-dual construction can be found at the end of \cref{sec:CKLagrangians}.

@@

Much of the complexity for $D$-dimensional vector theories at
loop-level arises due to the mixing between external polarization and
internal loop momenta. As noted, one advantage of expanding ansatz
construction towards functions of rational kinematics is the prospect
of stripping out overall factors of gauge invariant tensors. Rather
than building an ansatz from the irreducible scalar products of
\cref{eq:ym-basis},
\begin{equation}
N^{\text{YM}}_{m,L} = \sum_i \mathcal{O}_i \mathcal{R}^{(i)}_{m,L}
\end{equation}
where $\mathcal{O}_i$ are the basis tensors of
\cite{Bern:2017tuc,Carrasco:2019yyn}, and $ \mathcal{R}^{(i)}_{m,L}$ is a
rational function of scalar products.
